%!TEX root = tommaso-thesis.tex
%!TEX spellcheck = en_US

\begin{acknowledgements}

\begin{center}
\begin{minipage}{.93\linewidth}
``Oh, I've had such a curious dream!'' said Alice, and she told her
sister, as well as she could remember them, all these strange Adventures
of hers that you have just been reading about; and when she had finished, her sister kissed her, and said ``It \emph{was} a curious dream, dear, certainly: but now run in to your tea; it's getting late.''

So Alice got up and ran off, thinking while she ran, as well she might,
what a wonderful dream it had been.
\end{minipage}
\end{center}

\smallskip


The years during my PhD were, unsurprisingly, among the most intense in my life.
During these years my world changed---several times.
Most of all I became a father, the most astonishing change that I have experienced in my life.

Many crossed my path during these years and contributed in shaping this experience.
I thank you all for the path you shared with me: I hope I was able to give you at least a fraction of what I received.

Thanks to Michele, advisor and friend.
I know I have been an unconventional PhD student, if anything because no PhD student is a conventional one.
Nonetheless, you had the patience to wait for me during weird times.
Maybe not crazy-Rumanian weird, but still weird times.
Yours was an example that left a mark in me.
I was no Mike Ross, but I still I hope that my passing left a trace in the REVEAL family.

Speaking of which, a big thank goes to the REVEALers, in particular those who were, for me, its \emph{hard core} members: Andrea, Luca, and Roberto, who shared with me this journey from the beginning.
I loved you. And I hated you. Often both at the same time.
I find amazing how much we changed together during these years.
You are family to me.

Thanks to the new generation REVEALers who shared with me this last part of my journey: Bin, Emad, Jevgenija, Csaba, and Gabriele.
You brought a fresh touch into the group and it was a lot of fun.
I wish you the best of luck with your adventure.

Thanks to the Dean's office, for accepting my madness and backing it up with even more madness.
Oh, yes, and for the chocolate, too.

Thanks to the Tymchuk family, Yuriy, Natalia, and Sophia.
You literally made us part of your family.
I know that our bond will last in time, regardless of how far apart we will be.

Thanks to the Pharo community, who allowed me to bring ShoreLine to life.
I had to almost implement my own rest server ---several times--- to do that but, hey, it was fun.

Thanks to Alberto Bacchelli, who brought me to Lugano in the first place and occasionally acted as ghostwriter.
I owe you a lot for inviting me during my master's thesis.
We share so many inside jokes that I am not able to pick one. \emph{Fottutorato}.

Thanks to Marco D'Ambros, who took part with me in a lot of funny adventures during my early days in Lugano, even if he remembers none of them.

Thanks to Giulia \emph{Pizzinato}, for being on our side when we needed it.
I'm happy that Samuele had the oportunity of knowing you and witnessing some sparks of crazy happiness.
There will always be pizza for you at our home.

A special mention goes to the people in charge of making badges at conferences: Rarely I have seen such a creativity as the one that you used in writing my name.
I have to admit that ``Sasso, Tommaso Dal'' is a masterpiece.

Finally, a deep thanks goes to my family who, directly or indirectly, lived this adventure with me.
Mom and Dad, you supported me in ways that I just started to understand, now that I am a father.
I am proud of the name I bring: You made me part of two families who share a silent, powerful bond.

Samuele, my son, you brought the light in our lives.
Looking at you exploring the world reminds me of how fulfilling it is to improve a little every day.
I will always be proud of you.

Elena, you have been a constant presence at my side.
We have been through a lot and I am proud of what we have done.
I know that we have wonderful things waiting for us.
\\It will be amazing.


\begin{flushright}
So Long, and Thanks for All the Fish\\
Tommaso
\end{flushright}
%We gratefully acknowledge the financial support of the Swiss National Science foundation for the project ``HI-SEA'' (SNF Project No. 146734).

\end{acknowledgements}
