%!TEX root = ../tommaso-thesis.tex
\chapter{List of Publications}\label{ch:publications}


\publication{Content Classification of Development Emails}{Alberto Bacchelli, Tommaso Dal Sasso, Marco D’Ambros, Michele Lanza}{In Proceedings of ICSE 2012 (34th International Conference on Software Engineering), pp. 375--385, IEEE CS Press, 2012.}{Bacc2012a}

\publication{A closer look at bugs}{Tommaso Dal Sasso, Michele Lanza}{In Proceedings of VISSOFT 2013 (1st IEEE Working Conference on Software Visualization), pp. 1--4, IEEE CS Press, 2013.}{DalS2013a}

\publication{in*Bug: Visual analytics of bug repositories}{Tommaso Dal Sasso, Michele Lanza}{In Proceedings of CSMR-WCRE 2014 (1st Joint Meeeting of the European Conference on Software Maintenance and Reengineering and the Working Conference on Reverse Engineering), pages 415--419, IEEE CS Press, 2014.}{DalS2014a}

\publication{Managing Software Defects}{Tommaso Dal Sasso}{In Proceedings of ICSME 2014 (30th International Conference on Software Maintenance and Evolution), page 669, Doctoral Symposium, 2014.}{DalS2014b}

\publication{Misery Loves Company: CrowdStacking Traces to Aid Problem Detection.}{Tommaso Dal Sasso, Andrea Mocci, Michele Lanza}{In Proceedings of SANER 2015 (22nd IEEE International Conference on Software Analysis, Evolution, and Reengineering), pp. 131--140, 2015.}{DalS2015a}

\publication{Blended, Not Stirred: Multiconcern visualization of large software systems}{Tommaso Dal Sasso, Roberto Minelli, Andrea Mocci, Michele Lanza}{In Proceedings of VISSOFT 2015 (3rd IEEE Working Conference on Software Visualization), pp. 106--115 2015.}{DalS2015b}

\publication{What Makes a Satisficing Bug Report}{Tommaso Dal Sasso, Andrea Mocci, Michele Lanza}{In Proceedings of QRS 2016 (The 2016 IEEE International Conference on Quality, Reliability, and Security), pp. 164--174, IEEE CS Press, 2016.}{DalS2016a}

\publication{How to Gamify Software Engineering}{Tommaso Dal Sasso, Andrea Mocci, Michele Lanza, Ebrisa Mastrodicasa}{In Proceedings of SANER 2017 (24th IEEE International Conference on Software Analysis, Evolution, and Reengineering), pp. 261--271, IEEE CS Press, 2017.}{DalS2017a}

\publication{Sympathy for the Devil: Reified Collection of Runtime Errors}{Tommaso Dal Sasso, Andrei Chis, Andrea Mocci, Tudor Girba, Michele Lanza}{In Proceedings of PLATEAU 2017 (8th International Workshop on Evaluation and Usability of Programming Languages and Tools), pp. 1--8, ACM Press, 2017}{DalS2017b}
